\begin{center}\textbf{Acknowledgements}
\end{center}

First and foremost, I would like to thank my advisor Jeffrey Teo for his mentorship, support, and kindness. Jeffrey gave me the freedom to pursue problems and projects of my choice. He exposed me to numerous intellectually stimulating projects and has been extremely patient and generous with his time in teaching me. His breadth of mathematics and physics knowledge and the ability to teach advanced topics in a clear and thorough manner has benefitted me tremendously. 

I am thankful for his constant mentorship in writing papers and preparing talks, which has improved my communication and writing skills drastically. I would also like to thank Jeffrey for giving me numerous opportunities to travel to many conferences, summer schools, and workshops, which helped me immensely in my physics career and were great learning experiences. I could not have asked for a better advisor and will always remain in awe of his genius and kindness. 

I would like to thank Gia-Wei Chern for the valuable discussions and for his help and support when I was applying for jobs. He is one of the easiest persons to talk to and has always been generous with his time and research ideas. I would like to thank Israel Klich whose courses helped me prepare for physics research and who has also been part of my research committee for the last four years. His questions and feedback during my research talks have been really helpful. I would also like to thank Prof. Craig Dukes for being the chair of my research committee, and for his advice and encouragement during my research committee talks. I am grateful to Thomas Mark for being a part of my thesis committee.

I enjoyed collaborating and discussing physics with a number of graduate students in the physics department, I would especially like to thank Alex Sirota, Meng Hua, Sharmistha Sahoo and Jing Luo for various collaborations. I am thankful to Peter Arnold, Diana Vaman and Michael Fowler for being wonderful and inspiring teachers. I am also grateful to Maksim Bychkov, whom I worked for over many semesters teaching physics labs. 

I would like to thank my parents, Saiqa and Dr. Humayun for their constant love and support, and for enabling me to pursue my dreams from a very young age. Lastly, I would like to thank my wife, Saman Hussain who has been a source of constant strength and inspiration and has been extremely supportive throughout my Ph.D.


	
